### 2021-01-20 introduction.tex
Update introduction.tex

\begin{document}
\section{Introduction}
Session types are a type formalism which can be used to specify and verify communication protocols between two or more communicating agents~\cite{honda93,takeuchihonda94,hondavasconcelos98,carbonehonda07}.
\todo{These two settings are very different. The $\pi$-calculus is a mobile process algebra preceding inside which session types were developed and GV is an ad-hoc calculus made to demonstrate how propositions can be used as types.}
Session types have been studied for many programming paradigms, including concurrent and functional paradigms. Most notably, they have been defined for the $\pi$-calculus~\cite{sangiorgiwalker01}---a process calculus for communication and concurrency, and for Good Variation (GV)~\cite{wadler15,lindleymorris15}---a linear concurrent $\lambda$-calculus.

Session-typed concurrent $\lambda$-calculus and $\pi$-calculus satisfy communication safety and session fidelity, namely that communication proceeds without any type mismatch of the data being transmitted and follows the predefined protocol specified by a session type.
Another key property in a concurrent setting is deadlock freedom, stating that agents communicate without getting stuck in cyclic dependencies. Unfortunately, deadlock freedom is a more involving property and session types alone are not enough to rule out deadlocks in cyclic structures of multiple interleaved sessions. Several techniques have been developed to address deadlock freedom in the $\pi$-calculus and concurrent $\lambda$-calculus with session types.

@@

With regards to GV, the original work on session types in GV~\cite{gayvasconcelos10,gayvasconcelos12} did not satisfy deadlock freedom. Later on, calculi in the GV family~\cite{wadler15,lindleymorris15} have achieved deadlock freedom via a syntactic restriction, \ie by combining channel creation and thread spawning into a single operation, called fork. The fork construct is the term representation of the cut construct in CP, which evaluates to a channel restriction combined with a parallel composition. Unfortunately, as with CP, this is restrictive as it limits communication structures to trees. There are several benefits of working with a functional language as opposed to name-passing calculi. Functional languages support higher-order functions, and have a capability for abstraction not usually present in process calculi. Working within a functional language allows to derive extensions of the communication capabilities of the language via well-understood extensions of the functional fragment, \ie deriving internal/external choice via sum types. Finally, in concurrent functional calculi, there is a clear separation between a program that the user writes, and a configuration, which is the state of a system as it evaluates. In process calculi, these are conflated.
Moreover, the benefit of working with GV over other session-typed functional languages is that GV has strong ties to linear logic, via its relation to CP~\cite{wadler14}, and consequently it has strong formal properties, \eg deadlock freedom, albeit in a restrictive way.
Following the above observations, we aim to further investigate GV and deadlock freedom, and thus we pose our research question:

@@

 \item
 \textbf{Solving GV problems}. PGV addresses several problems in the original GV language, most notably:
 	\begin{enumerate}
	\item our version does not have the pseudo-type $S^\sharp$;
	\item our structural congruence is type preserving.
	\end{enumerate}
\item
\textbf{Updated Priority CP.} We present an updated version of Priority CP~\cite{dardhagay18} in \cref{sec:pcp}. We remove commuting conversions and move away from reduction as cut elimination, towards reduction based on structural congruence, as in process calculi. We present Milner's Cyclic Scheduler in PCP (\cref{sec:pcp-example}).

\item
\textbf{Connection to linear logic}. The connection of PGV to logic is given via a sound and complete encoding of PCP to PGV in~\cref{sec:pcp} and prove that it preserves typing (\cref{thm:pcp-to-pgv-confs-preservation}) and satisfies operational correspondence (\cref{thm:pcp-to-pgv-operational-correspondence-soundness,thm:pcp-to-pgv-operational-correspondence-completeness}).  We showcase it with the encoding of cyclic scheduler to PGV (\cref{sec:pgv-example}).
\end{enumerate}
\end{document}

>>>>

\begin{document}
\section{Introduction}
Session types are a type formalism which can be used to specify and verify communication protocols between two or more communicating agents~\cite{honda93,takeuchihonda94,hondavasconcelos98,carbonehonda07}.
Session types have been studied for many programming paradigms, most notably for concurrent and functional paradigms. They have been defined for the $\pi$-calculus~\cite{sangiorgiwalker01}---a process calculus for communication and concurrency, and for Good Variation (GV)~\cite{wadler15,lindleymorris15}---a linear concurrent $\lambda$-calculus.

Session type systems guarantee communication safety and session fidelity, namely that communication proceeds without any type mismatch of the data being transmitted and follows the predefined protocol specified by a session type.
Another key property in a concurrent setting is deadlock freedom, stating that agents communicate without getting stuck in cyclic dependencies. Unfortunately, deadlock freedom is a more involving property and session types alone are not enough to rule out deadlocks in cyclic structures of multiple interleaved sessions. Several techniques have been developed to address deadlock freedom in the $\pi$-calculus and concurrent $\lambda$-calculus with session types.

@@

With regards to GV, the original work on session types in GV~\cite{gayvasconcelos10,gayvasconcelos12} did not satisfy deadlock freedom. Later on, calculi in the GV family~\cite{wadler15,lindleymorris15} have achieved deadlock freedom via a syntactic restriction, \ie by combining channel creation and thread spawning into a single operation, called fork. The fork construct is the term representation of the cut construct in CP, which evaluates to a channel restriction combined with a parallel composition. Unfortunately, as with CP, this is restrictive as it limits communication structures to trees. There are several benefits of working with a functional language as opposed to name-passing calculi. Functional languages support higher-order functions, and have a capability for abstraction not usually present in process calculi. Working within a functional language allows to derive extensions of the communication capabilities of the language via well-understood extensions of the functional fragment, \ie deriving internal/external choice via sum types. Finally, in concurrent functional calculi, there is a clear separation between a program that the user writes, and a configuration, which is the state of a system as it evaluates. In process calculi, these are conflated.
Moreover, the benefit of working with GV over other session-typed functional languages is that GV has strong ties to linear logic, via its relation to CP~\cite{wadler14}, and consequently it has strong formal properties, \eg deadlock freedom, albeit in a restrictive way.

Following the above observations, we aim to further investigate GV and deadlock freedom, and thus we pose our research question.

@@

 \item
 \textbf{Solving GV problems}. PGV addresses several problems in the original GV language, most notably:
 	\begin{enumerate}
	\item our version does not have the pseudo-type $S^\sharp$;
	\item our structural congruence is type preserving.
	\end{enumerate}
\item
\textbf{Updated Priority CP.} We present an updated version of Priority CP~\cite{dardhagay18} in \cref{sec:pcp}. We remove commuting conversions and move away from reduction as cut elimination, towards reduction based on structural congruence, as it is standard in process calculi. %We present Milner's Cyclic Scheduler in PCP (\cref{sec:pcp-example}).

\item
\textbf{Connection to linear logic}. The connection of PGV to logic is given via a \emph{sound and complete encoding} of PCP to PGV in~\cref{sec:pcp} and prove that it preserves typing (\cref{thm:pcp-to-pgv-confs-preservation}) and satisfies operational correspondence (\cref{thm:pcp-to-pgv-operational-correspondence-soundness,thm:pcp-to-pgv-operational-correspondence-completeness}).  %We showcase it with the encoding of cyclic scheduler to PGV (\cref{sec:pgv-example}).
\end{enumerate}
\end{document}
